---
title: "Statistical Process Control"
output:
  pdf_document: default
  html_document: default
date: "2025-04-08"
---

## Executive Summary
This report presents a Statistical Process Control (SPC) analysis of brewery operations based on data from 100 brewing batches across 8 beer styles and 10 production locations. The analysis focuses on identifying process variations, control patterns, and quality correlations to optimize brewing operations.

### Key findings:

All key process variables are in statistical control but show suspicious run patterns
pH levels and gravity values demonstrate strong consistency across beer styles
Losses during production vary significantly across different beer styles and production stages
Production location plays a significant role in brewhouse efficiency variation
Temperature and pH level show weak but measurable correlations with quality scores

## Introduction
Statistical Process Control (SPC) is a method for monitoring, controlling, and improving a process through statistical analysis. In brewing operations, SPC helps maintain product consistency, identify sources of variation, and prevent quality issues before they reach customers.
This analysis examines brewery data using SPC principles to understand:

1. Process stability and capability
2. Factors affecting quality
3. Variations across beer styles and production locations
4. Opportunities for process improvement

## Process Stability Analysis
### Control Limits for Key Process Variables
We calculated 3-sigma control limits for key process variables:

### Process Control Findings

1. Point Conformance: All process variables showed no points outside the 3-sigma control limits, which indicates good statistical control of the process.
2. Run Patterns: Despite point conformance, run tests revealed suspicious patterns in all variables:

- pH Level: Longest run of 14 consecutive points above/below mean
- Gravity: Longest run of 10 consecutive points
- Temperature: Longest run of 10 consecutive points
- Other variables: Similar patterns with runs of 8-16 consecutive points

3. Coefficient of Variation (CV):

Process variables show varying degrees of consistency:
Gravity is the most consistent (CV = 1.38%)
pH Level shows good consistency (CV = 5.55%)
Temperature and Alcohol Content show moderate variation
Loss measurements show high variability (CV > 37%)


```{r echo=FALSE}
library(qcc)
library(readr)
library(dplyr)

# Read the data
beer_data <- read_csv("bottle_dataset.csv")

# Set seed for reproducibility
set.seed(42)

# Group data for X-bar and R charts (creating subgroups of size 5)
# First, let's sort by Beer_Style to make meaningful subgroups
beer_data <- beer_data %>% arrange(Beer_Style)

# Function to create subgroups
create_subgroups <- function(data, variable, subgroup_size = 5) {
  values <- data[[variable]]
  n <- length(values)
  # Trim to make it divisible by subgroup_size
  n_groups <- floor(n/subgroup_size)
  values <- values[1:(n_groups*subgroup_size)]
  # Reshape into matrix with subgroup_size columns
  matrix(values, ncol = subgroup_size, byrow = TRUE)
}

# X-bar and R charts for pH Level
#pH_subgroups <- create_subgroups(beer_data, "pH_Level")
#pH_xbar <- qcc(pH_subgroups, type = "xbar", title = "X-bar Chart for pH Level")
#pH_R <- qcc(pH_subgroups, type = "R", title = "R Chart for pH Level")

# X-bar and R charts for other key variables
# variables <- c("Gravity", "Alcohol_Content", "Temperature")
variables <- c("Quality_Score")
for(var in variables) {
  var_subgroups <- create_subgroups(beer_data, var)
  var_xbar <- qcc(var_subgroups, type = "xbar", title = paste("X-bar Chart for", var))
  var_R <- qcc(var_subgroups, type = "R", title = paste("R Chart for", var))
  # Print the charts
  print(var_xbar)
  print(var_R)
}

# Generate defect and defective data for attribute charts
# For a brewery, defects could be related to:
# - Underfilled bottles (defective unit)
# - Off-flavors (defects per unit)
# - Label issues (defects per unit)
# - Cap sealing problems (defective unit)

# Sample size (inspection units)
n <- nrow(beer_data)
sample_size <- rep(100, n) # Assuming each row represents a batch with 100 inspections

# Generate defective counts for np and p charts
# Bottles with any defect (binary: defective or not)
# Probability of defect related to Quality_Score (lower score, higher probability)
defect_prob <- 0.1 * (10 - beer_data$Quality_Score)/10
defectives <- rbinom(n, sample_size, defect_prob)

# Generate defect counts for u charts
# Number of defects (multiple defects possible per unit)
# Rate related to both Quality_Score and Loss_During_Bottling_Kegging
defect_rate <- 0.2 * (10 - beer_data$Quality_Score)/10 + 
               0.3 * beer_data$Loss_During_Bottling_Kegging
defects <- rpois(n, defect_rate * 100) # Multiply by 100 for the sample size

# np chart (number of defective units)
np_chart <- qcc(defectives, type = "np", sizes = sample_size, title = "np Chart for Defective Bottles")

# p chart (proportion of defective units)
p_chart <- qcc(defectives, type = "p", sizes = sample_size, title = "p Chart for Proportion of Defective Bottles")

# u chart (defects per unit)
u_chart <- qcc(defects, type = "u", sizes = sample_size, title = "u Chart for Defects per Bottle")

# Run charts by beer style with attribute data
styles <- unique(beer_data$Beer_Style)
for(i in 1:length(styles)) {
  style <- styles[i]
  style_indices <- which(beer_data$Beer_Style == style)
  
  if(length(style_indices) >= 5) { # Only if we have enough data
    # x-bar and R charts for this style
    # style_pH <- create_subgroups(beer_data[style_indices,], "pH_Level")
    # if(nrow(style_pH) > 1) {
    #   style_xbar <- qcc(style_pH, type = "xbar", title = paste("X-bar Chart for pH Level -", style))
    #   style_R <- qcc(style_pH, type = "R", title = paste("R Chart for pH Level -", style))
    # }
    
    # p chart for this style
    style_defectives <- defectives[style_indices]
    style_sizes <- sample_size[style_indices]
    if(length(style_defectives) > 1) {
      style_p <- qcc(style_defectives, type = "p", sizes = style_sizes, 
                    title = paste("p Chart for Defective Rate -", style))
    }
  }
}

# Location-based analysis with attribute data
locations <- unique(beer_data$Location)
for(loc in locations) {
  loc_indices <- which(beer_data$Location == loc)
  
  if(length(loc_indices) >= 5) { # Only if we have enough data
    # u chart for this location
    loc_defects <- defects[loc_indices]
    loc_sizes <- sample_size[loc_indices]
    if(length(loc_defects) > 1) {
      loc_u <- qcc(loc_defects, type = "u", sizes = loc_sizes,
                  title = paste("u Chart for Defect Rate -", loc))
    }
  }
}
```

```{r}
# Capability analysis for Quality_Score
Quality_capability <- process.capability(
  qcc(beer_data$Quality_Score, type = "xbar.one"),
  spec.limits = c(7, 10)  # Example specification limits
)
```

